\textbf{The Mint Scale: A Fresh Validation of the Multimodal Interoception Questionnaire and Comparison to the MAIA, BPQ and IAS}

### 2025-08-19 paper/manuscript.tex
final

\abstract{TO DO. }

\keywords{Interoception questionnaire, interoceptive accuracy scale,
MAIA, Mint Validation, Body Awareness}

\authornote{\par{\addORCIDlink{Dominique
Makowski}{0000-0001-5375-9967}}\par{\addORCIDlink{Ana
Neves}{0009-0006-0020-7599}}\par{\addORCIDlink{Giulia
Poreiro}{0000-0002-2343-5109}} 
\par{ }
\par{     \begin{tcolorbox}[enhanced jigsaw, left=2mm, arc=.35mm, bottomrule=.15mm, leftrule=.75mm, rightrule=.15mm, colback=white, breakable, opacityback=0, colframe=quarto-callout-note-color-frame, toprule=.15mm]

This preprint is a non-peer-reviewed work from the
\href{https://realitybending.github.io/}{\textbf{Reality Bending Lab}}.
\begin{center}
\includegraphics[width=0.2\linewidth,height=\textheight,keepaspectratio]{manuscript_files/mediabag/ReBeL_LogoOnly_hu114.png}
\end{center}

@@

Main issues in existing questionnaires:

\begin{itemize}
\tightlist
\item
  Either heavily based on theories (e.g., focusing on specific
  ``alleged'' dimensions), despite shaky evidence for said-theories
\item
  Do not control for context (which leads to variability in
  interpretation and occurence)
\item
  Often quite narrow in the modalities covered (or scattered, like the
  IAS)
\end{itemize}

@@

TODO: write intro.

\begin{itemize}
\tightlist
\item
  Goal of study 1: to generate a lot of items, analyze its structure and
  reduce them to a balanced set of items.
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Possible (although not at all significant) bias consistent across
  indices in favour of a greater similarity with the modality-grouping
  condition. Futher research with more data (here we only had
  \textasciitilde200 observation per condition) might be able to yield
  more conclusive results. In general, testing the presentation mode is
  a novel approach which should be more widely applied to psychometric
  validations.
\item
  We grouped some of the dimensions under ``Awareness'', despite being
  similar to items included in the IAS typically conceptualized as
  ``Accuracy''. However, we believe that subjective scales cannot be
  pure measures of accuracy, as they are inherently subjective and thus
  reflect a degree of awareness (of one's accurate noticing or
  sensation).
\item
  Olfactory compensation: pseudo-interoceptive pathway whereby people
  with interoceptive deficits would compensate by gathering evidence
  about their internal state externally (e.g., via smell).
\item
  The metacluster ``Sensitivity'' seems to be grouping iterms related to
  visceropcetion and (hyper)sensitivity (Derm).
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Goal of study 2: to validate the Mint scale against other
  interoception scales.
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Structure: Low stability of \emph{Olfa} as belonging in the deficit
  group. More studies are needed to investigate its place and cluster
  invariance: it is possible that its association might depend on other
  factors or categories.
\item
  Main contender is MAIA. But its most sensitive dimensions seem to be
  the least related to interoception, which is dangerous as it might
  lead to findings with associations misattributed to ``Interoception''
  (just because it correlate with some MAIA dimensions), but it might
  have nothing to do with interoception per se and more with emotion
  regulation.
\item
  Basically the pattern is clear: everything high-level, metacognitive
  and highly subjective related to wellbeing and general affective state
  is the most strongly related to the MAIA through some of its Emotion
  regulation dimensions. All the other outcomes actually related to
  interoceptive processes are most strongly predicted by the Mint.
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Future directions: needs to test relationship with physiological
  measures of interoception.
\item
  Also, the structure might still require a bit of deeper investigation,
  in particular with some of its facets (\emph{Olfa} and \emph{Derm})
  which might display different associations depending on the context or
  the population.
\end{itemize}

>>>>

\abstract{Interoception, the sensing of the body's internal state, is an
important component of homeostatic and allostatic regulation with
wide-ranging implications for mental and physical health, yet its
measurement is hindered by tools with conceptual and psychometric
issues. To address these limitations, we developed and validated the
Multimodal Interoception Questionnaire (Mint). In Study 1 (N=559), a
comprehensive item pool using a novel ``modality-by-context'' framework
was generated and then reduced using psychometric network analysis. In
Study 2 (N=737), we validated the Mint in an independent sample,
establishing its structure, comparing its predictive power against
established interoception questionnaires (MAIA-2, BPQ, IAS), and across
a wide range of clinical and cognitive outcomes. This rigorous process
yielded a 33-item scale with a stable hierarchical structure comprising
three metaclusters (Interoceptive Deficit, Interoceptive Awareness,
Visceroception) and 11 distinct facets. The Mint demonstrated superior
criterion validity, consistently outperforming existing scales in
predicting outcomes tied to altered bodily perception, including
alexithymia, ADHD, autism, and somatic symptom clusters. The Mint offers
a psychometrically robust and nuanced measure of self-reported
interoception that integrates and extends existing scales, providing
researchers and clinicians with a practical and readily usable tool to
investigate the role of bodily sensations in health and disease. }

\keywords{Interoception questionnaire, interoceptive accuracy scale,
MAIA, Mint Validation, Body Awareness, Health}

\authornote{\par{\addORCIDlink{Dominique
Makowski}{0000-0001-5375-9967}}\par{\addORCIDlink{Ana
Neves}{0009-0006-0020-7599}}\par{\addORCIDlink{Giulia
Poreiro}{0000-0002-2343-5109}} 
\par{ }
\par{     \begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, opacityback=0, colback=white, breakable, toprule=.15mm, arc=.35mm, left=2mm, rightrule=.15mm, bottomrule=.15mm, colframe=quarto-callout-note-color-frame]

This preprint is a non-peer-reviewed work from the
\href{https://realitybending.github.io/}{\textbf{Reality Bending Lab}}.
\includegraphics[width=0.2\linewidth,height=\textheight,keepaspectratio]{manuscript_files/mediabag/ReBeL_LogoOnly_hu114.png}

@@

Interoception - the sensing and interpretation of the body's internal
physiological signals - plays a fundamental role in processes ranging
from homeostatic regulation to emotion, decision-making, and mental
health (\citeproc{ref-craig2002how}{Craig, 2002};
\citeproc{ref-critchley2017interoception}{Critchley \& Garfinkel,
2017}). Despite its growing recognition, the field has been hindered by
inconsistent definitions and fragmented measurement approaches
(\citeproc{ref-desmedt2022measures}{Desmedt et al., 2022},
\citeproc{ref-desmedt2025discrepancies}{2025};
\citeproc{ref-khalsa2018interoception}{Khalsa et al., 2018}). Available
tools include ``objective'' physiological tasks (e.g., heartbeat
counting or tracking) and ``subjective'' self-reports, and encompass
both ``explicit'' paradigms that direct participants' attention to
internal sensations and ``implicit'' paradigms that do not (e.g.,
heartbeat evoked potentials recorded during rest). These assessments are
often targeted at only one of several interoceptive modalities (e.g.,
cardioception, respiroception, gastroception), and are mapped onto
theoretical dimensions (e.g., accuracy, sensitivity, awareness,
\citeproc{ref-garfinkel2015knowing}{Garfinkel et al., 2015}) within a
rapidly evolving theoretical landscape. However, no consensus exists on
a gold-standard measure witheach approach having distinct limitations.
For example, physiological tasks often require specialised equipment and
expertise, are time-consuming, and are challenging to administer at
scale, whereas questionnaires are constrained by their inherent
subjectivity. Crucially, the psychometric quality of many of these
measures is questionable (e.g., see criticism about the heartbeat
counting task, \citeproc{ref-zamariola2018interoceptive}{Zamariola et
al., 2018}), yielding low inter-task correlations
(\citeproc{ref-murphy2018classifying}{Murphy et al., 2018}) and raising
important concerns regarding their reliability and validity (e.g.,
measures presented as measuring one theoretical dimension but likely
measuring something else). This lack of conceptual clarity and
measurement convergence impedes scientific progress in comparing,
replicating, and interpreting interoception-related findings,
emphasising the need to develop improved tools that align with modern
psychometric standards.

The present work focuses on self-report questionnaires, which offer a
scalable and accessible tool complementing physiological tasks. Among
the numerous interoceptive questionnaires available, a handful dominate
the literature (\citeproc{ref-desmedt2022measures}{Desmedt et al.,
2022}), including the Body Perception Questionnaire (BPQ,
\citeproc{ref-kolacz2018body}{Kolacz et al., 2018}), the
Multidimensional Assessment of Interoceptive Awareness (MAIA,
\citeproc{ref-mehling2018multidimensional}{Mehling et al., 2018}), and
the Interoceptive Accuracy Scale (IAS,
\citeproc{ref-murphy2018classifying}{Murphy et al., 2018}). These tools
were developed in different eras of a rapidly evolving field, each with
its own objectives and philosophies. For example, the BPQ, one of the
earliest measures, captures bodily awareness and autonomic nervous
system (ANS) reactivity by focusing on experiences of physiological
stress reactions in organs innervated by the ANS, such as those
triggered by stress or trauma (\citeproc{ref-kolacz2018body}{Kolacz et
al., 2018}). The MAIA incorporated ideas from the then-booming emotion
regulation and mindfulness fields, including subscales assessing
adaptive coping strategies and metacognitive insights about the body. In
contrast, the IAS was recently developed to circumvent these high-level
- possibly tangential - constructs, instead capturing subjective reports
of concrete bodily events (e.g., coughing, sneezing) as proxies for
interoceptive accuracy.

These differing philosophies have practical consequences, leading to
poor convergence between scales that are assumed to measure the same
construct (\citeproc{ref-vig2022questionnaires}{Vig et al., 2022}) - a
classic example of the ``jingle fallacy'' where the shared label of
``interoception'' can mask underlying measurand differences, hindering
replication and theoretical integration. In acknowledgement of this
issue, efforts have been made to align subjective measures with
theoretical models that outline different facets of interoception (e.g.,
accuracy, attention, deficits). However, focusing on developing measures
that are theory-driven may however be premature when theoretical models
are in flux, and when a topographic mapping to self-report dimensions is
not guaranteed. For example, interoceptive distinctions that are
relevant at a neuroanatomical or cognitive level might not necessarily
map onto the structure of subjective experiences. More fundamentally,
existing interoception questionnaires typically fail to consider the
full range of interoceptive modalities and the context in which
interoceptive experiences occur. Most existing questionnaires focus on a
narrow range of bodily modalities, with a heavy bias towards
cardio-respiratory sensations at the expense of others like gastric,
urogenital, or thermoregulatory signals. Perhaps most critically,
existing measures ask participants to make general assessments across
most/all contexts, which does not capture the potential homeostatic
relevance of attending or detecting internal signals. Without a
contextual reference point, there is likely to be substantial
variability in how participants interpret and respond to items. For
example, a general item about ``feeling your heart beat'' is ambiguous
since its salience, frequency, experience and interpretation would
differ depending on the context, e.g., on whether it occurs during
physical exercise, a moment of anxiety, or quiet rest (e.g., someone who
exercises regularly, or is anxious, might more easily recall an instance
of occurrence of feeling their heart beating and respond based on that
contextual saliency).

To address these gaps, we developed and validated the Multimodal
Interoception Questionnaire (Mint), a new questionnaire that
systematically addresses the above limitations. In Study 1, we describe
the bottom-up development of the Mint, using a ``modality-by-context''
framework to generate a comprehensive item pool, which was then refined
using cutting-edge psychometric network analysis. In Study 2, we
validated the final 33-item version of the Mint against the established
interoception questionnaires, demonstrating its robust structure and
predictive validity for a range of clinically-relevant outcomes.

@@

The aim of Study 1 was to develop a new interoception questionnaire by
generating a comprehensive item pool and subsequently refining it into a
psychometrically robust scale. To overcome the limitations of existing
questionnaires, we adopted a bottom-up, theoretically agnostic approach.
Rather than imposing a pre-defined theoretical structure on
questionnaire items, we instead focused on two tangible and
systematically controllable aspects of interoceptive experience: the
\emph{modality} of the sensation and the \emph{context} in which it
occurs. Our objective was not to develop a questionnaire measuring a new
or theoretically distinct facet, but instead build on existing research
and synthesise existing ideas and frameworks into a comprehensive and
modern tool that captures both the richness of interoceptive experiences
and the most prominent facets of interoception.

Item generation was guided by a ``modality-by-context'' grid. For
\emph{modalities}, we expanded beyond the typical cardio-respiratory
focus to encompass a wide range of bodily systems: cardiac, respiratory,
gastric, genital, bladder \& colon, and skin \& temperature (the latter
of which being increasingly recognized as a key interoceptive channel,
\citeproc{ref-crucianelli2023role}{Crucianelli \& Ehrsson, 2023}). We
also included a ``general state'' category to capture the integrated
sense of the body's overall physiological condition
(\citeproc{ref-craig2002how}{Craig, 2002};
\citeproc{ref-craig2003interoception}{Craig, 2003}). For
\emph{contexts}, we designed facets to capture specific situational
contexts like negative (anxious) and positive (sexual) arousal, which
reflect distinct bodily states with differing interoceptive signatures.
Facets were also designed to capture other experiential qualities
tapping into elaboration and threshold of perception, such as
sensitivity, accuracy, and confusion. This systematic framework yielded
a diverse pool of 120 initial items, ensuring comprehensive coverage
while minimising the ambiguity inherent in context-free questions.

The initial pool of 120 was then reduced using cutting-edge psychometric
network-based techniques, identifying and removing redundant items and
revealing the most stable and coherent dimensional structure. We also
designed the study to enable an examination of structural invariance
depending on presentation format (e.g., how items are grouped together,
whether by modality, facet or randomly) which is an often overlooked
psychometric quality. This rigorous, multi-stage process was designed to
produce a subset of items with a clear, stable, and theoretically rich
hierarchical structure, ready for formal validation in Study 2.

@@

The aim of Study 1 was to develop and refine a large, systematically
generated pool of interoceptive items into a psychometrically sound and
structurally coherent questionnaire. Starting with 120 items derived
from a modality-by-context grid, we employed a multi-step, data-driven
reduction process using cutting-edge structure analysis methods. This
procedure yielded a final, balanced set of 45 items, organised within a
stable hierarchical structure comprising three higher-order dimensions
(Interoceptive Deficits, Interoceptive Awareness, Interoceptive
Sensitivity) and 12 distinct lower-level facets.

The resulting hierarchical structure offers novel yet interpretable
insights into self-reported interoception, with flexible degrees of
granularity. Notably, we categorised several facets (typically related
to objective external behaviours, like coughing or sneezing) under the
label ``Awareness''. Although other questionnaires might refer to this
as ``Accuracy'', we preferred to preserve the notion that subjective
self-reports, by their very nature, cannot be a pure measure of
objective accuracy and are, at best, a proxy for subjective certainty or
awareness of interoceptive processes. Furthermore, our bottom-up
approach revealed intriguing facets, such as ``Olfactory Compensation''.
This finding provides evidence for a pseudo-interoceptive pathway,
whereby individuals with poorer access to internal signals may
compensate by relying on external cues (e.g., odours) to infer their
bodily state. The ``Sensitivity'' metacluster also grouped core
visceroceptive facets (e.g., ``Cardioception'', ``Gastroception'') with
``Dermal Hypersensitivity'', pointing to a potential common dimension of
heightened perceptual sensitivity to bodily sensations.

An additional innovation of Study 1 was examining the effect of item
presentation on the resulting psychometric structure by comparing
structures across different grouping conditions. Although there were no
significant effects of item presentation, a consistent (though
non-significant) trend suggested that grouping items by modality (rather
than by facet or randomly) could be more conducive to revealing the
underlying psychometric structure. This finding, while preliminary due
to the sample size per condition (\(N \pm 200\)), highlights a
potentially important and under-investigated aspect of questionnaire
design worthy of future research.

@@

Study 1 successfully leveraged a systematic item generation process and
modern analytical techniques to produce a 45-item interoception
questionnaire with a clear, stable, and theoretically rich hierarchical
structure. Study 2 then sought to formally validate the Mint scale by
examining its psychometric properties and its relationship with existing
measures in a large, independent sample. To this end, we sought to 1)
re-assess the structure of the shorter Mint version and 2) establish its
convergent validity by examining correlations with other established and
popular interoception questionnaires (namely the MAIA-2, BPQ, and IAS),
as well as other related measures (e.g., alexithymia). Finally, we
assessed the Mint's criterion validity by directly comparing the
predictive power of all interoception questionnaires across a broad
range of outcomes, including emotional, cognitive, and clinical
variables.

@@

Study 2 sought to formally validate the Mint in a large, independent
sample, a crucial step for establishing the psychometric soundness and
unique contribution of the scale for interoception research. This
involved re-assessing its structure in a new sample and establishing its
convergent and criterion validity against other existing interoception
questionnaires. The analysis led to a final, refined 33-item Mint scale
with a stable and theoretically coherent hierarchical structure, which
demonstrated substantial predictive power across a range of outcomes.

Structure re-validation is an important step following item reduction
since reducing a large item pool can alter the perceived meaning and
relationships between the remaining items. For instance, a questionnaire
with many items suggesting abnormality or pathology might influence the
interpretation of otherwise neutral items. While the original structure
of the Mint was mostly replicated, the re-validation facilitated the
clarification of several higher-order dimensions (e.g., the
``Sensitivity'' metacluster became more clearly defined around core
interoceptive clusters). The final Mint structure, comprising 11 facets
within 3 metaclusters, yielded excellent structural consistency and
stability. The \emph{Olfa} (Olfactory Compensation) cluster had lower
stability within the \emph{Deficit} metacluster, possibly indicative of
a more complex role. For instance, this cluster could be related to
adaptive compensatory processes in some participants and maladaptive
strategies in others. Future studies should further investigate its
place and invariance across different populations and contexts.

The criterion validity analysis revealed a clear and compelling pattern
of results. The Mint scale, particularly its \emph{Deficit} metacluster,
consistently outperformed existing interoceptive scales in predicting
outcomes directly related to alterations in bodily experience and
self-regulation, including alexithymia, dissociative experiences, ADHD,
autism, and various somatic symptoms. The MAIA-2 showed an advantage
over the Mint for predicting outcomes related to general well-being
(e.g., life satisfaction, depression, and anxiety). However, its
predictive superiority was almost entirely driven by dimensions related
to emotion regulation abilities, namely \emph{Trusting} and
\emph{Not-Worrying}, whose inclusion as core interoceptive facets have
been questioned (\citeproc{ref-ferentzi2021examining}{Ferentzi et al.,
2021}). While valuable, the MAIA-2's most predictive subscales appear
less related to core interoceptive processing and more to metacognitive
beliefs about one's body and emotional coping. The Mint, in contrast,
appears to tap more directly into distinct facets of interoceptive
experience, from visceral sensitivity to perceived deficits. Indeed,
when the emotion-regulation dimensions were removed (in the \emph{iMAIA}
version), the MAIA-2's predictive power for well-being outcomes dropped
considerably, often below that of the Mint. Overall, for high-level,
metacognitive aspects of affective state and well-being, the MAIA-2's
emotion regulation facets are highly informative. For outcomes tied more
specifically to bodily processes, clinical conditions characterised by
sensory dysregulation, and nuanced interoceptive experiences, the Mint
offers superior predictive validity and specificity.

@@

Across two studies, we developed and validated the Multimodal
Interoception Questionnaire (Mint), a new 33-item self-report measure of
interoception. Our approach addressed key limitations of existing
questionnaires by starting with a systematically generated, broad item
pool that explicitly accounted for context and assessed a wide range of
interoceptive modalities. The Mint is a psychometrically robust
instrument with a stable, hierarchical structure comprising three
higher-order metaclusters, namely \emph{Interoceptive Deficit},
\emph{Interoceptive Awareness}, and \emph{Visceroception}, and 11
distinct lower-level facets. Study 2 validated the Mint, demonstrating
that it not only captures variance present in other popular scales
(MAIA-2, BPQ and IAS) but also offers superior predictive power for a
range of clinical and cognitive outcomes, particularly those tied to
altered bodily perception.

A key advantage of the Mint is its potential ability to disentangle core
interoceptive experiences from more general emotion regulation
abilities, an identified problem in interoception research where
findings may be misattributed to ``interoception'' when they are driven
by emotion regulation capacities
(\citeproc{ref-murphy2020testing}{Murphy et al., 2020}). Our findings
show that although the MAIA-2 is a strong predictor of general
well-being, this predictive power is driven by its \emph{Trusting} and
\emph{Not-Worrying} subscales, which heavily tap into emotion regulation
skills and coping strategies. The Mint, in contrast, taps more directly
into interoceptive components, and has excellent predictive power for
conditions where alterations in bodily sensing are a core feature,
including alexithymia, ADHD, autism, and somatic symptoms. Overall, the
Mint provides a more specific measure of self-reported interoceptive
processing across contexts and modalities, offering researchers a
granular and comprehensive tool with which to further understand the
contribution of bodily awareness to human health and disease.

We note several limitations that could serve as the impetus for future
research efforts. First, as a self-report measure, the Mint captures
subjective beliefs about interoceptive abilities, not objective
accuracy. A natural next step will be to establish the extent to which
subjective reports on the Mint correspond to objective interoceptive
performance (e.g., heartbeat tracking tasks, heartbeat evoked
potentials). Second, future research could systematically examine the
usefulness of shorter versions of the Mint, for example, by
investigating the tradeoffs incurred by versions with only the one or
two most central facets per metacluster on psychometric qualities and
predictive power.

While the overall structure of the Mint is stable, some facets warrant
deeper investigation. The \emph{Olfa} (Olfactory Compensation) and
\emph{Derm} (Dermal Hypersensitivity) clusters, for instance, showed
more complex patterns of association. Their placement within the
\emph{Deficit} metacluster is theoretically intriguing since the former
aligns with predictive coding models of pseudo-interoception (or
extero-interoception) where individuals rely on exteroceptive cues (like
odours) to infer internal states
(\citeproc{ref-paulus2010interoception}{Paulus \& Stein, 2010}), while
the latter links visceral and cutaneous sensitivity, possibly reflecting
a general trait of heightened bodily awareness or a shared mechanism of
central sensitization, observed in conditions like chronic pain
(\citeproc{ref-di2016alexithymia}{Di Tella \& Castelli, 2016}). Future
studies should examine the invariance of the Mint's structure across
different populations (e.g., clinical \emph{vs.} non-clinical groups)
and contexts (e.g., heightened arousal \emph{vs.} rest) to clarify the
role of these more nuanced facets. This work could reveal whether the
relevance and pattern of associations are shared or dissociable across
conditions like anxiety disorders, chronic pain, and neurodevelopmental
disorders.

In conclusion, the Mint scale represents a fresh take on interoception
questionnaires and a significant methodological advancement for the
field of interoception research. By providing a validated, structurally
coherent, and nuanced tool, it allows for a more precise assessment of
self-reported interoceptive experiences, paving the way for a deeper
understanding of the multifaceted role of interoception in human
psychology and pathophysiology.
